```{r}
library(DBI)
db = dbConnect(RSQLite::SQLite(), dbname  = "class_2023-10-27/employees.sqlite")
```

## Exercise 1

### 1. The total costs in payroll for this company

```{sql, connection=db}
SELECT SUM(salary) AS total_payroll FROM employees;
```

### 2. The average salary within each department

```{sql, connection=db}
SELECT dept, AVG(salary) AS avg_salary FROM employees GROUP BY dept;
```


## Exercise 2

### Lets try to create a table that has a new column - abv_avg which 
### contains how much more (or less) than the average, for their
### department, each person is paid.

```{sql, connection=db}
SELECT *, salary - avg_salary AS diff  FROM
  employees LEFT JOIN
  ( SELECT dept, AVG(salary) AS avg_salary FROM employees GROUP BY dept )
  USING (dept)
  ORDER BY dept, diff DESC
;
```



## Exercise 3
### Write a query that determines the total number of seats available on all of the planes that flew out of New York in 2013.

```{r}
db = dbConnect(RSQLite::SQLite(), dbname  = "class_2023-10-27/flights.sqlite")
```

### Wrong answer
```{sql, connection=db}
SELECT SUM(seats) FROM flights NATURAL LEFT JOIN planes;
```

### Correct answer

```{sql, connection=db}
SELECT SUM(seats) FROM flights LEFT JOIN planes USING (tailnum);
```
